---
title: "Data Visualization"
author: "Meagan Brown"
date: '2022-08-17'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)

# Load libraries
library(tidyverse)
library(here)
library(janitor)

library(patchwork)
library(ggrepel)
library(gghighlight)
library(paletteer)
library(ggExtra)
library(ggbeeswarm)

library(gapminder)

library(sf)
```

### Read in lizard data

```{r}
lizards <- read_csv(here("data_tidy", "lizards.csv"))
```

### Make a heatmap
```{r}
lizard_counts <- lizards %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  count(year = lubridate::year(date),
        common_name) %>% 
  drop_na()
```

```{r}
ggplot(data = lizard_counts,
       aes(x = year,
           y = common_name)) +
  geom_tile(aes(fill = n), 
            show.legend = FALSE) +
  geom_text(aes(label = n), color = "white", size = 3) +
  scale_fill_gradientn(colors = c("navy", "red", "orange")) +
  theme_minimal()
```

### Beeswarm plot
```{r}
whiptails <- lizards %>% 
  filter(common_name == "western whiptail") 
```

```{r}
ggplot(data = whiptails,
       aes(x = sex,
           y = weight)) +
  geom_beeswarm(size = 0.5) +
  geom_boxplot(fill = NA)
  theme_minimal() 
```

### Marginal Plot
```{r}
whiptails <- lizards %>% 
  filter(common_name == "western whiptail") %>% 
  drop_na(total_length, weight) # will only drop if in this specific row/column is NA
```

```{r}
whiptail_plot <- ggplot(data = whiptails,
       aes(x = total_length,
           y = weight)) +
  geom_point(aes(color = sex)) +
  theme(legend.position = "bottom") +
  theme_classic()

ggMarginal(whiptail_plot, type = "boxplot", 
           groupColour = TRUE)
```

### Patchwork
```{r}
whiptail_bee <- ggplot(data = whiptails,
       aes(x = sex,
           y = weight)) +
  geom_beeswarm(size = 0.5) +
  geom_boxplot(fill = NA)
  theme_minimal()
  
  whiptail_plot <- ggplot(data = whiptails,
       aes(x = total_length,
           y = weight)) +
  geom_point(aes(color = sex)) +
  theme(legend.position = "bottom") +
  theme_classic() 
```

```{r}
(whiptail_bee * whiptail_plot) &
  theme_classic()
```


```{r}
ggplot(data = lizards,
       aes(x = total_length,
           y = weight)) +
  geom_point()
```

Create a histogram of total lengths for all lizards in the data set

```{r}
ggplot(data = lizards,
       aes(x = total_length)) +
  geom_histogram(color = "orange",
                 fill = "purple",
                 size = 0.3,
                 linetype = "dotted")
```

```{r}
ggplot(data = lizards,
       aes(x = total_length,
           y = weight)) +
  geom_point(shape = 22, 
             color = "cyan4",
             fill = "yellow",
             size = 4,
             alpha = 0.5)
```

```{r}
ggplot(data = lizards,
       aes(x = total_length,
           y = weight)) +
  geom_point(aes(color = common_name,
                 size = total_length),
             alpha = 0.5) +
  theme_light() +
  facet_wrap(~common_name,
             scales = "free") #each scale is different
```

```{r}
lizard_count <- lizards %>% 
  group_by(common_name) %>% 
  summarize(counts = n())
#OR
lizard_counts <- lizards %>% 
  count(common_name)
```

```{r}
#Column graph that plots value n()
ggplot(data = lizard_count,
       aes(y = fct_reorder(common_name, counts), #orders based on value instead of alphabetical
           x = counts)) +
  geom_col()
```

Create a scatterplot facteted by species for total length versus weight. The color and shape of points should differ based on species (common_name). Change the theme. Make each faceted panel have flexible axis limits (scales should be free).

```{r}
ggplot(data = lizards,
       aes(x = total_length,
           y = weight)) +
  geom_point(aes(color = common_name,
                 shape = common_name),
             size = 3,
             alpha = 0.6
             ) +
  facet_wrap(~common_name, scales = "free") +
  theme_light() +
  labs(x = "Total length (mm)",
       y = "Weight (g)",
       title = "Jornada Basin lizard sizes",
       subtitle = "All lizards follow the standard length weight relatiionship",
       caption = "Data: Jornada Basin LTER, CCBy")
```

### Fun with scales

```{r}
ggplot(data = lizards,
       aes(x = total_length,
           y = weight)) +
  geom_point(aes(color = weight)) +
  scale_colour_gradient(low = "blue",
                        high = "red")
```

```{r}
ggplot(data = lizards,
       aes(x = total_length,
           y = weight)) +
  geom_point(aes(color = weight)) +
  scale_colour_steps(low = "red",
                     high = "black")
```

```{r}
ggplot(data = lizards,
       aes(x = total_length,
           y = weight)) +
  geom_point(aes(color = total_length)) +
  scale_color_steps2(low = "red",
                     mid = "white",
                     high = "orange",
                     midpoint = 150, 
                     breaks = c(50, 75, 150, 180, 220, 280))
```

```{r}
ggplot(data = lizards,
       aes(x = total_length,
           y = weight)) +
  geom_point(aes(color = total_length)) +
  scale_color_stepsn(colors = c("orange", "red", "purple"),
                     breaks = seq(from = 0, 
                                  to = 400,
                                  by = 60))
```

Make a horizontal boxplot with common name on the y-axis, total length on the x-axis, with boxplot fill chnaging based on the common name

```{r}
lizards_median <- lizards %>% 
  mutate(common_name = fct_reorder(common_name,
                                   total_length,
                                   .fun = median)) #reorder boxplot by median value 

ggplot(data = lizards_median,
       aes(y = common_name,
           x = total_length)) +
  geom_boxplot(aes(fill = common_name), show.legend = FALSE) +
  scale_fill_paletteer_d(palette = "metbrew::Archambault") +
  theme_minimal()
```

```{r}
ggplot(data = lizards,
       aes(x = total_length,
           y = weight)) +
  geom_point(aes(color = weight)) +
  theme(panel.grid.major.x = element_line(color = "white"), # make x grid lines white
        panel.grid.minor.y = element_blank(), #remove y grid lines
        axis.title.x = element_text(color = "purple"), # change x axis text color
        axis.text.y = element_text(color = "green"), #change y axis text color
        text = element_text(size = 18), 
        panel.background = element_rect(color = "purple",
                                        fill = "white")) +
  annotate("text", x = 300, y = 50, label = "Woooooo", color = "magenta") +
  geom_vline(xintercept = 250,
             linetype = "dashed",
             color = "orange",
             size = 4)
```

Repulsive lables with `ggrepel`
```{r}
wws_lizards <- lizards %>% 
  filter(common_name == "western whiptail",
         site == "sand")

ggplot(data = wws_lizards,
       aes(x = total_length,
           y = weight)) +
  geom_point() +
  geom_text_repel(aes(label = toe_num)) #off centers labels dodges

gapminder %>% 
  filter(year == 2002, 
         continent == "Europe") %>% 
  ggplot(aes(x = gdpPercap,
             y = lifeExp)) +
  geom_point() +
  geom_text_repel(aes(label = country), size = 3)

ggplot(data = lizards,
       aes(x = total_length,
           y = weight)) +
  geom_point() +
  gghighlight(weight > 30,
              label_key = toe_num)
```


### Make a map
```{r}
jornada_veg <- read_sf(here("data_raw", "spatial_vegetation", "doc.kml"))

ggplot() +
  geom_sf(data = jornada_veg,
          aes(fill = Name),
          color = NA) +
  theme_void() +
  scale_fill_paletteer_d(palette = "ggthemes::manyeys")
```

```{r}
library(tidyverse)
library(palmerpenguins)

ggplot(data = na.omit(penguins), aes(x = species, y = body_mass_g, shape = sex)) +
  geom_point(alpha = 0.2, position = position_jitterdodge(dodge.width = 0.8)) + 
  # means & standard devs
  stat_summary(mapping = aes(color = species), 
               fun = "mean", geom = "point", size = 4,
               position = position_dodge(width = 0.8)) +
  stat_summary(mapping = aes(color = species),
               fun = "mean", geom = "errorbar", size = 1, width = 0.2,
               fun.max = function(x) mean(x) + sd(x),
               fun.min = function(x) mean(x) - sd(x),
               position = position_dodge(width = 0.8)) +
  # change colors/shapes
  scale_color_manual(values = c("#FF8C02", "#A93FF1", "#148F90"), name = "Species") + 
  scale_shape_manual(values = c(15, 16), name = "Sex") + 
  # add nicer axis labels + title + caption
  labs(x = "Penguin Species", 
       y = "Body Mass (g)",
       title = "Body mass of female vs. male adelie, chinstrap, and gentoo penguins",
       subtitle = "Colored points represent mean body mass (± SD)",
       caption = "Data Source: Dr. Kristen Gorman, LTER Palmer Station") +
  theme_classic() + 
  theme(
    plot.title = element_text(hjust = 0, size = 14),
    axis.text = element_text(color = "black", size = 10),
    axis.title = element_text(color = "black", size = 14),
    plot.caption = element_text(size = 7, hjust = 1, color = "gray", face = "italic"),
    panel.border = element_rect(color = "black", size = 0.7, fill = NA))


```

